Report on the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-04-09 WSSSPE4_report/wg_best_practices_sustainability.tex
Misc edits

Many open source projects for research software document their best practices that contributed to the sustainability of the software.  Many open source projects also document software metrics they use to define their research software project's success.  This project endeavors to aggregate several sources of these best practices into a consolidated list of best practices, as and to also do the same for software metrics in another consolidated list.  Our team will then create a workflow to evaluate how open source projects stand up to these lists.

>>>>

Many open source projects for research software document their best practices that contributed to the sustainability of the software.  Many open source projects also document software metrics they use to define their research software project's success.  This project endeavors to aggregate several sources of these best practices into a consolidated list of best practices, and similar for software metrics.  Our team will then create a workflow to evaluate how open source projects stand up to these lists.

### 2017-04-09 WSSSPE4_report/wg_open_research_index.tex
Misc edits

The aim of this group is to investigate the building of an index of research products in an open sustainable manner.  Our goal is not to eliminate commercial products, but to build on what?s there and provide data and services that are missing.

@@

\item Iterate use case with stakeholders (who, when),
e.g., funders, Researchfish, researchers, ?

>>>>

The aim of this group is to investigate the building of an index of research products in an open sustainable manner.  Our goal is not to eliminate commercial products, but to build on what's there and provide data and services that are missing.

@@

\item Iterate use case with stakeholders (who, when),
e.g., funders, Researchfish, researchers ?

### 2017-04-09 WSSSPE4_report/wg_prototyping.tex
Misc edits

%from Santiago
There is a productivity bottleneck---yet to be solved---in HPC from the human
perspective, first identified in the 1980s~\cite{barstow1982automatic}. Of all
domain-specific scientists, only a percentage of those use simulations due to
perceptions mostly related to the difficulty of using and developing those tools.
XSEDE and other resources have become a gateway for successfully increasing the
basis of HPC scientific users by facilitating their access to infrastructure and
tools~\cite{towns2014xsede}, but the development challenge remains unaddressed.
Of all scientists who use simulations, only a small percentage know how
to develop prototypes or full applications that can be latter scaled by computer
scientists and engineers. Writing new parallel applications is unusual for
domain scientists and thus has created a complex dependency on specialized
scientific programmers, who are scarce~\cite{post2005computational}. However,
 the capacity to quickly envision and prototype applications by domain experts
 ---later to be fully implemented into by scientific programmers---is critical for
 acquiring maturity and proficiency in the grand challenges the Exascale
 attempts to solve, and moreover for allowing scientists to develop their own
 ideas quickly and productively~\cite{vinter2015prototyping}. The latter, in
 addition, would increase service utilization time in HPC systems, a critical
measure of efficiency. Considering the existing code base in scientific
computing, software infrastructures that allow a clean transition from legacy
systems to new Exascale platforms while preserving flexible prototyping
 towards the future are central~\cite{hwu2015transitioning}.

>>>>

%from Santiago
There is a productivity bottleneck---yet to be solved---in HPC from the human
perspective, first identified in the 1980s~\cite{barstow1982automatic}. Of all
domain-specific scientists, only a percentage of those use simulations due to
perceptions mostly related to the difficulty of using and developing those tools.
XSEDE and other resources have become a gateway for successfully increasing the
basis of HPC scientific users by facilitating their access to infrastructure and
tools~\cite{towns2014xsede}, but the development challenge remains unaddressed.
Of all scientists who use simulations, only a small percentage know how
to develop prototypes or full applications that can be latter scaled by computer
scientists and engineers. Writing new parallel applications is unusual for
domain scientists and thus has created a complex dependency on specialized
scientific programmers, who are scarce~\cite{post2005computational}. However,
 the capacity to quickly envision and prototype applications by domain experts
 ---later to be fully implemented into by scientific programmers---is critical for
 acquiring maturity and proficiency in the grand challenges that Exascale
 attempts to solve, and moreover for allowing scientists to develop their own
 ideas quickly and productively~\cite{vinter2015prototyping}. The latter, in
 addition, would increase service utilization time in HPC systems, a critical
measure of efficiency. Considering the existing code base in scientific
computing, software infrastructures that allow a clean transition from legacy
systems to new Exascale platforms while preserving flexible prototyping
 towards the future are central~\cite{hwu2015transitioning}.
